\chapter{Proposal Alignment and Scope Narrowing}
\label{app:proposal_alignment}

This appendix makes the proposal-to-submission mapping explicit. The purpose is not to inflate delivered scope after the fact, but to show which promised items were completed, which were narrowed, and what repository or thesis evidence supports that judgment.

\small
\setlength{\LTleft}{0pt}
\setlength{\LTright}{0pt}
\begin{longtable}{p{0.23\textwidth}p{0.31\textwidth}p{0.13\textwidth}p{0.25\textwidth}}
\caption{Proposal alignment and scope narrowing for the final submission.}\label{tab:proposal_alignment}\\
\toprule
\textbf{Proposal task} & \textbf{Evidence} & \textbf{Status} & \textbf{Submission wording} \\
\midrule
\endfirsthead
\toprule
\textbf{Proposal task} & \textbf{Evidence} & \textbf{Status} & \textbf{Submission wording} \\
\midrule
\endhead
\bottomrule
\endfoot
Unified dataset and preprocessing pipeline & \texttt{src/data/io.py}, \texttt{scripts/dataset\_utils.py}, Chapter~\ref{ch:method} & Delivered & The submission provides a canonical JSONL schema, prompt/context assembly, validation checks, and documented preprocessing controls. \\
Design/model establishment & Technical attachments, Chapter~\ref{ch:method}, Chapter~\ref{ch:system_demo} & Delivered & The architecture, module map, data flow, and ML pipeline are documented as final submission artifacts. \\
Rules baseline & \texttt{src/baselines/rules.py}, \texttt{src/llm\_jailbreak\_detector/rules\_detector.py}, demo documentation & Delivered & A deterministic offline rules baseline is implemented and used as the guaranteed demo path. \\
Learned detector / LoRA pipeline & \texttt{scripts/train\_lora.py}, \texttt{src/llm\_jailbreak\_detector/lora\_detector.py}, \texttt{runs/week7\_norm\_only/config.json} & Delivered & The project delivers a local-artifact LoRA classifier with persisted validation threshold and reproducible run metadata. \\
Robustness evaluation and locked pack & \texttt{reports/week7/locked\_eval\_pack/week7\_norm\_only/}, Chapters~\ref{ch:experiments}--\ref{ch:robustness} & Delivered & Unicode, adv2, and rewrite stress results are packaged in the locked Week~7 evidence bundle and integrated into the thesis. \\
Comparison against existing guardrails & Scope-narrowing note in the submission bundle, Chapter~\ref{ch:limitations_ethics} & Scope narrowed; external benchmark not delivered & The final submission narrows this deliverable to an offline reproducible detector study because no runnable local artifact for proposal-named external guardrails was available in the submission environment. \\
Installable / runnable software & \texttt{pyproject.toml}, \texttt{src/llm\_jailbreak\_detector/}, demo bundle & Delivered & The repository ships an installable package and an examiner-safe rules-only demo flow. \\
Testing documentation & \texttt{docs/Testing\_Documentation.md}, CI workflow, \texttt{tests/} & Delivered & Automated and manual verification procedures are documented, with local results and CI coverage recorded. \\
User manual & \texttt{docs/User\_Manual.md}, school submission bundle & Delivered & The submission includes final install, command, schema, demo, and troubleshooting guidance. \\
Technical attachment pack & \texttt{submission/Technical Attachments/} & Delivered & The school-required technical attachments are bundled under school-facing names in one obvious folder. \\
\end{longtable}
\normalsize
