\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

Large language model applications often assemble prompts from multiple text sources --- user input, retrieved passages, and attached context --- which makes prompt-layer attacks such as jailbreaks and prompt injection a practical deployment risk. This thesis studies prompt-layer guardrail detection as a binary classification problem over a single text payload. The delivered system is an offline-capable detector with a command-line interface and Python library, exposing an always-available rules baseline and an optional learned LoRA classifier that runs strictly from local artifacts.

Methodologically, the thesis makes a threshold-transfer evaluation protocol central to the analysis. The learned detector is implemented as a parameter-efficient LoRA sequence classifier with a DeBERTa-v3-base backbone, and the final selected run is \texttt{week7\_norm\_only}. A decision threshold of \(\tau = 0.7340\) is calibrated on validation at target false-positive rate (FPR) \(= 1\%\) and then transferred unchanged to held-out splits, avoiding optimistic per-split threshold tuning. Robustness is assessed using reproducible perturbation families that probe common evasion channels: Unicode obfuscation, character-level noise (adv2), and paraphrase-like rewriting.

Empirically, the learned detector achieves strong ranking quality on the main evaluation distribution (\(\mathrm{AUROC}=0.9958\) on \texttt{test\_main}), but the fixed operating point does not transfer stably under shift. At the same transferred threshold, benign FPR rises from \(0.0094\) on validation to \(0.0311\) on \texttt{test\_main}, \(0.5816\) on \texttt{test\_jbb}, and \(0.6403\) on \texttt{test\_main\_adv2}. The main contribution is therefore not a claim of deployment-ready robustness, but an examiner-verifiable demonstration that high AUROC and AUPRC do not guarantee low-FPR deployment behavior under distribution shift. The thesis argues that the threshold should be treated as a monitored policy object, that the rules baseline remains a conservative offline fallback, and that the committed Week~7 locked evaluation pack should be treated as the authoritative source of reported counts, metrics, and figures.

\noindent\textbf{Keywords:} large language models, prompt injection, jailbreak detection, guardrails, threshold calibration, robustness evaluation
